---
title: "Life Insurance Simulation"
author: "Andy Merlino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

We provide whole life insurance.  

How much money do we need to set aside today to have a certain probability that we can payout the entire life insurance  benefit to all insurees?

```{r load_stuff}
library(insuree)
library(knitr) # for tables

options(scipen=999) # turn off scientific notation
```

Let's create some life insurance policies to insure.

```{r create_policies, results = "asis"}
policies <- data.frame(id = c("A001", "A002", "A003", "A004", "A005", "A006"),
                       gender = c("male", "female", "female", "male", "male", "female"),
                       age = c(40, 51, 70, 45, 65, 35),
                       death_benefit = c(200000, 500000, 100000, 500000, 1000000, 2000000))

knitr::kable(policies)
```

As you can see, we have male and female insurees of varying ages.  They also have varying death benefits.

We are going to create an object of class `insuree::Insuree` to represent each one of our insurees.  Each of these `insuree::Insuree` objects is defined by an actual life table.  We will use the official 2010 US social security life table as provided on the [Official Social Security Website](http://www.ssa.gov/oact/STATS/table4c6.html).  I named this life table `qx_table` and it comes with the package.

```{r life_tables}
# select interest rate
rate <- 0.04
# select number of times to run simulation
n <- 1000

# create male life table from qx_data (provided with the package)
qx_male <- insuree::LifeTable(x = qx_data$x, q_x = qx_data$male_qx)
qx_male <- insuree::ActuarialTable(i = rep(rate, times = length(qx_male@x)), qx_male)

# create female life table from qx_data (provided with the package)
qx_female <- insuree::LifeTable(x = qx_data$x, q_x = qx_data$female_qx)
qx_female <- insuree::ActuarialTable(i = rep(rate, times = length(qx_female@x)), qx_female)
```

We simulate each individual's death in accordance with the multinomial distribution paramaterized by the probabilties from either the male or female qx table.  We run `r n` observations of each individual's future life expectancy.  We assume each individual future life expectancy is independent and identically distributed.  We discount the death benefit, assuming a `r rate * 100` % interest rate, for the number of years the insuree survives.  All death benefits are paid at the end of the year of death.

First lets assign each insuree to a `insuree::Insuree` object.

```{r}
# use lifetable package to define each policy as a Insuree object
max_m <- max(qx_male@x)
max_f <- max(qx_female@x)

hold <- list()
for (j in 1:nrow(policies)) {
  hold[[j]] <- if (policies$gender[j] == "male") {
    insuree::Insuree(x_ = policies$age[j], 
                     t_ = max_m - policies$age[j],
                     benefit = rep(policies$death_benefit[j], max_m - policies$age[j]), 
                     qx_male)
  } else {
    insuree::Insuree(x_ = policies$age[j], 
                     t_ = max_f - policies$age[j],
                     benefit = rep(policies$death_benefit[j], max_f - policies$age[j]), 
                     qx_female)
  }
}
```

Now that we have assigned each individual to a `insuree::Insuree` object we can run the simulation.  The present value of an `insuree::Insuree` object can be simulated with the `insuree:rpv` method. 

```{r, result = "asis"}
# run simulation on each insuree
# we are only interested in the present value of the simulation here
out <- lapply(hold, function(k) insuree::rpv(k, n = n)$pv)
out <- matrix(unlist(out), ncol = n, byrow = TRUE)
total <- apply(out, 2, sum)
quantiles <- quantile(total, seq(0.70, 0.95, by = 0.05))
sim_mean <- mean(total)

# make it look decent for printing
out <- data.frame("Risk Level" = c("mean", names(quantiles)),
                 "Value" = format(c(sim_mean, quantiles), big.mark = ","))
knitr::kable(out, row.names = FALSE)
```

For additional information on how the simulation is run please see the `insuree::rdeath` function.