---
title: "Life Insurance Simulation"
author: "Andy Merlino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


We provide whole life insurance.  

How much money do we need to set aside today to have a certain probability that we can payout the entire life insurance  benefit to all insurees?

```{r load_stuff}
library(lifetable)
library(knitr)

options(scipen=999) # turn off scientific notation
```

Let's create some life insurance policies to insure.

```{r create_policies, results = "asis"}
policies <- data.frame(id = c("A001", "A002", "A003", "A004", "A005", "A006"),
                       gender = c("male", "female", "female", "male", "male", "female"),
                       age = c(40, 51, 70, 45, 65, 35),
                       death_benefit = c(200000, 500000, 100000, 500000, 1000000, 2000000))

knitr::kable(policies)
```

As you can see, we have male and female insurees of varying ages.  They also have varying death benefits.

We are going to create an object of class *lifetable::Z_x* to represent each one of our insurees.  Each of these *lifetable::Z_x* objects is defined by an actual life table.  We will use the official 2010 US social security life table as provided on the [Official Social Security Website](http://www.ssa.gov/oact/STATS/table4c6.html).  I named this life table *qx_table* and it comes with the package.

```{r life_tables}
# select interest rate
rate <- 0.04
# select number of times to run simulation
n <- 1000

# create male life table from qx_data (provided with the package)
qx_male <- lifetable::LifeTable(x = qx_data$x, q_x = qx_data$male_qx)
qx_male <- lifetable::ActuarialTable(i = rep(rate, times = length(qx_male@x)), qx_male)

# create female life table from qx_data (provided with the package)
qx_female <- lifetable::LifeTable(x = qx_data$x, q_x = qx_data$female_qx)
qx_female <- lifetable::ActuarialTable(i = rep(rate, times = length(qx_female@x)), qx_female)
```

We simulate each individual's death in accordance with the multinomial distribution paramaterized by the probabilties from either the male or female qx table.  We run `r n` observations of each individual's future life expectancy.  We assume each individual future life expectancy is independent and identically distributed.  We discount the death benefit, assuming a `r rate * 100` % interest rate, for the number of years the insuree survives.  All death benefits are paid at the end of the year of death.

```{r}
# use lifetable package to define each policy as a Z_x object
hold <- list()
for (j in 1:nrow(policies)) {
  hold[[j]] <- if (policies$gender[j] == "male") {
    lifetable::Z_x(x_ = policies$age[j], 
                    t_ = max(qx_male@x) - policies$age[j],
                    benefit = policies$death_benefit[j], 
                    qx_male)
  } else {
    lifetable::Z_x(x_ = policies$age[j], 
                    t_ = max(qx_female@x) - policies$age[j],
                    benefit = policies$death_benefit[j], 
                    qx_female)
  }
}
```

Now that we have assigned each individual to a *lifetable::Z_x* object we can run the simulation.  The time and present value of a *lifetable::Z_x* object can be simulated with the *lifetable:rdeath* method. 

```{r, result = "asis"}
# run simulation on each insuree
# we are only interested in the present value of the simulation here
out <- lapply(hold, function(k) lifetable::rdeath(k, n = n)$pv)
out <- matrix(unlist(out), ncol = n, byrow = TRUE)
total <- apply(out, 2, sum)
quantiles <- quantile(total, seq(0.70, 0.95, by = 0.05))
sim_mean <- mean(total)

# make it look decent for printing
out <- data.frame("Risk Level" = c("mean", names(quantiles)),
                 "Value" = format(c(sim_mean, quantiles), big.mark = ","))
knitr::kable(out, row.names = FALSE)
```

For additional information on how the simulation is run please see the lifetable::rdeath function.
